---
title: "iMove checks"
author: "James McMahon"
date: "13/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# source("extract-data/00_setup-environment.R")
library(gt)
library(scales)

imove_clean <- read_rds(path(server_dir, "..", "final_data", "IMOVE_data.rds")) %>%
  as_tibble() %>%
  mutate(across(
    c(
      idcountry,
      sex,
      hospitalward,
      icu,
      multiple_episode,
      multiple_hosp,
      prevhosp,
      hcw,
      smoking,
      pregnant,
      trimester,
      postpartum,
      residence,
      abdopain,
      ageusia,
      anosmia,
      chest,
      chills,
      confusion,
      conjunct,
      coryza,
      cough,
      dermato,
      diarr,
      dizzy,
      fever,
      general_deter,
      headache,
      malaise,
      myalgia,
      nausea,
      palp,
      sob,
      sorethroat,
      suddenonset,
      tach,
      vomit,
      abo,
      ct_res,
      ct_us_ecg,
      cxr,
      ecg_qt,
      ox_nasal,
      nebu,
      prone,
      trialdrugs,
      study_convpl,
      study_oth,
      vent,
      vent_type,
      closecont,
      closecont_type,
      anaemia,
      asplenia,
      asthma,
      cancer,
      dement,
      diabetes,
      heartdis,
      hypert,
      immuno,
      liverdis,
      lungdis,
      neuromusc,
      obese,
      rendis,
      rheumat,
      stroke,
      tuberc,
      outcome,
      deathcause,
      healthcare_contact,
      lab_covid,
      lab_covtest,
      lab_covtesttype,
      seq
    ),
    as_factor
  )) %>%
  mutate(across(c(age_m, los_hosp, los_icu), as.integer)) %>%

  # Add iso weeks and years
  mutate(across(where(is.Date), list(
    week = lubridate::isoweek,
    year = lubridate::isoyear
  ))) %>%

  # Days between symptom and swab
  mutate(swabdelay = time_length(onsetdate %--% swabdate, "days")) %>%

  # Creating a variable for presence of any chronic disease
  mutate(anychron = grepl(1, paste(liverdis, diabetes, heartdis, cancer, immuno, lungdis, rendis, dement, stroke, rheumat, anaemia, tuberc, asplenia, asthma, hypert, neuromusc, obese), fixed = TRUE)) %>%

  # Create age groups
  mutate(agegp = case_when(
    between(age_y, 0, 14) ~ 1L,
    between(age_y, 15, 49) ~ 2L,
    between(age_y, 50, 59) ~ 3L,
    between(age_y, 60, 69) ~ 4L,
    between(age_y, 70, 79) ~ 5L,
    between(age_y, 80, 89) ~ 6L,
    age_y >= 90 ~ 7L,
    TRUE ~ NA_integer_
  ))

q1 <- function(x, ...) quantile(x, 0.25, ...)
q3 <- function(x, ...) quantile(x, 0.75, ...)

n_lower_outliers <- function(x, ...) {
  x <- x[!is.na(x)]
  length(x[x < q1(x, ...)])
}

n_upper_outliers <- function(x, ...) {
  x <- x[!is.na(x)]
  length(x[x > q3(x, ...)])
}

n_missing <- function(x, ...) {
  sum(is.na(x), ...)
}
```

## Summary Checks

### Numeric checks

```{r numeric_checks}
imove_clean %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), list(
    missing = n_missing,
    min = min,
    max = max,
    mean = mean,
    q1 = q1,
    median = median,
    q3 = q3,
    sd = sd,
    lowoutlier = n_lower_outliers,
    highoutlier = n_upper_outliers
  ),
  na.rm = TRUE
  )) %>%
  pivot_longer(everything(),
    names_to = c("var", ".value"),
    names_pattern = "^(.+?(?:_.+?)?(?:_.+?)?(?:_.+?)?)_(\\w+?)$"
  ) %>%
  gt()
```

### Date checks

```{r date_checks}
imove_clean %>%
  select(where(is.Date)) %>%
  summarise(across(everything(), list(
    missing = n_missing,
    min = min,
    max = max,
    mean = mean,
    median = median
  ),
  na.rm = TRUE
  )) %>%
  pivot_longer(everything(),
    names_to = c("var", ".value"),
    names_pattern = "^(.+?(?:_.+?)?(?:_.+?)?(?:_.+?)?)_(\\w+?)$"
  ) %>%
  gt()
```



## Check covid case definitions

* Check not meeting case definition (=0)
* Check missing

```{r covid}

```

## Check all records have sex

**`r nrow(imove_clean) - sum(imove_clean$sex %in% c(0, 1))`** rows have missing sex (`r percent(1 - (sum(imove_clean$sex %in% c(0, 1)) / nrow(imove_clean)))`)

```{r sex}
imove_clean %>%
  count(sex) %>%
  gt()
```

## Check pregnancy

### Male pregnancy

Checking if any men are pregnant, have trimester or are post-partum.

```{r pregnant_male}
pregnant_male <- imove_clean %>%
  filter(
    sex == 1,
    pregnant == 1 | trimester != 8 | postpartum == 1
  ) %>%
  select(hosp_id2, pregnant, trimester, postpartum)

pregnant_male
```

Issues with **`r nrow(pregnant_male)`** records.

### Incorrect female pregnancy

Check for women who are pregnant and aged under 16 or over 50. Clearly not impossible.

```{r pregnant_female}
pregnant_female <- imove_clean %>%
  filter(
    sex == 0,
    pregnant == 1,
    age_y < 15 | age_y > 50
  ) %>%
  select(hosp_id2, age_y, pregnant, trimester, postpartum) %>%
  arrange(age_y)

pregnant_female
```

Possible issues with **`r nrow(pregnant_female)`** records.


## Check sari case definitions

* Check not meeting case definition (=0)
* Check missing

```{r sari}

```

## Check missing symptoms

Needs more work - need to do sari case definition and look for missing symptoms

```{r missing_sympt}
imove_clean <- imove_clean %>%
  mutate(
    resp_symptom = case_when(
      cough == 1 | sorethroat == 1 | sob == 1 ~ 1L,
      TRUE ~ 0L
    ),
    systemic_symptom = case_when(
      fever == 1 | malaise == 1 | headache == 1 | myalgia == 1 | general_deter == 1 ~ 1L,
      TRUE ~ 0L
    ),
    resp_covsymptom = case_when(
      cough == 1 | sorethroat == 1 | sob == 1 | coryza == 1 | tach == 1 ~ 1L,
      TRUE ~ 0L
    ),
    systemic_covsymptom = case_when(
      fever == 1 | chills == 1 | malaise == 1 | headache == 1 | myalgia == 1 | dermato == 1 | confusion == 1 | dizzy == 1 | general_deter == 1 ~ 1L,
      TRUE ~ 0L
    ),
    abdo_covsymptom = case_when(
      vomit == 1 | diarr == 1 | nausea == 1 | abdopain == 1 ~ 1L,
      TRUE ~ 0L
    ),
    other_covsymptom = case_when(
      anosmia == 1 | ageusia == 1 | chest == 1 | palp == 1 | conjunct == 1 ~ 1L,
      TRUE ~ 0L
    )
  )
```

## Check consent

This variable isn't in our data. Should it all be 1?

## Check antivirals

Don't have `antivir` variable

## Missing lab results

**`r imove_clean %>% filter(covid == 1, is.na(lab_covid)) %>% nrow()`** rows have positive COVID but no lab result (likely diagnosis data as COVID lab confirmed but no linked ECOSS record)

## Check age

**`r imove_clean %>% filter(is.na(age_y) & is.na(age_m)) %>% nrow()`** rows have missing ages. 
**`r imove_clean %>% filter(age_y < 0 | age_m < 0) %>% nrow()`** rows have negative ages. 

**`r imove_clean %>% filter(age_y > 105) %>% nrow()`** rows have an age > 105.

```{r age_hist}
ggplot(imove_clean, aes(x = age_y, fill = sex)) +
  geom_histogram(binwidth = 1) +
  theme_minimal() +
  scale_fill_brewer("Sex", palette = 3) +
  xlab("Age in years")
```

## Los checks

### Provided Los
**`r imove_clean %>% filter(los_hosp == time_length(admitdate %--% dischargedate, "days")) %>% nrow()`** rows have a provided hospital length of stay which is equal to the derived (days between admission and discharge dates).

**`r imove_clean %>% filter(los_icu == time_length(icuadmitdate %--% icudisdate, "days")) %>% nrow()`** rows have a provided ICU length of stay which is equal to the derived.


### Negative Los
**`r imove_clean %>% filter(los_hosp < 0) %>% nrow()`** rows have a negative provided hospital length of stay. 

**`r imove_clean %>% filter(los_icu < 0) %>% nrow()`** rows have a negative provided ICU length of stay. 

**`r imove_clean %>% filter(time_length(admitdate %--% dischargedate, "days") < 0) %>% nrow()`** rows have a negative derived hospital length of stay. 

**`r imove_clean %>% filter(time_length(icuadmitdate %--% icudisdate, "days") < 0) %>% nrow()`** rows have a negative derived ICU length of stay. 


### Extensive Los
**`r imove_clean %>% filter(los_hosp > 60) %>% nrow()`** rows have a provided hospital length of stay greater than 60 days. 

**`r imove_clean %>% filter(los_icu > 60) %>% nrow()`** rows have a provided ICU length of stay greater than 60 days.

**`r imove_clean %>% filter(time_length(admitdate %--% dischargedate, "days") > 60) %>% nrow()`** rows have a derived hospital length of stay greater than 60 days.

**`r imove_clean %>% filter(time_length(icuadmitdate %--% icudisdate, "days") > 60) %>% nrow()`** rows have a derived ICU length of stay greater than 60 days.

## ICU admission

**`r imove_clean %>% filter(is.na(icuadmitdate), icu == 1) %>% nrow()`** rows have a no ICU dates but are flagged with an ICU admission.

**`r imove_clean %>% filter(!is.na(icuadmitdate), icu == 0) %>% nrow()`** rows have a ICU dates but are flagged as having no ICU admission.

## Outcomes

### Deaths

**`r imove_clean %>% filter(is.na(deathdate), outcome == 1) %>% nrow()`** rows have no death-date but are flagged deceased.

**`r imove_clean %>% filter(!is.na(deathdate), outcome != 1 , !is.na(outcome)) %>% nrow()`** rows have a death-date but are not flagged as deceased.

**`r imove_clean %>% filter(!is.na(deathdate), is.na(outcome)) %>% nrow()`** rows have a death-date but have no outcome.

### Discharges

**`r imove_clean %>% filter(!is.na(dischargedate), outcome == 1) %>% nrow()`** rows have a discharge-date and are marked as deceased. - This seems like it is ok to me?

**`r imove_clean %>% filter(!is.na(dischargedate), is.na(outcome)) %>% nrow()`** rows have a discharge-date but have no outcome.

**`r imove_clean %>% filter(!is.na(dischargedate), outcome == 3) %>% nrow()`** rows have a discharge-date but are flagged as still hospitalised.

## GP visits

## Time to swab / onset
 
## Other dates

### Admissions before 15 March

### Inconsistent dates

### Dates outside of expected period

### All missing dates

## Chronic Conditions
