---
title: "Scotland Summary statistics"
author: "Alice Byres, Nicole Jarvie and James McMahon"
date: "`r Sys.Date() %>% format('%A, %d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "left"
)

library(tidyverse)
library(lubridate)
library(finalfit)
library(RCurl)
library(Hmisc)
library(ggplot2)
library(openxlsx)
library(gt)
library(ggplot2)


if (!exists("scot_data")) { # Use this with rmarkdown::render("scotland_summary.Rmd") for testing
  source("code/00 - Get Scottish data.R")
  source("code/01 - Prep data.R")
}
```

```{r data_prep, include = FALSE}
# Fixes -------------------------------------------------------------------
# Change the age groupings
scot_data <- scot_data %>%
  mutate(
    age.factor = case_when(
      age < 17 ~ "<17",
      age < 30 ~ "17-29",
      age < 40 ~ "30-39",
      age < 50 ~ "40-49",
      age < 60 ~ "50-59",
      age < 70 ~ "60-69",
      age < 80 ~ "70-79",
      is.na(age) ~ NA_character_,
      TRUE ~ "80+"
    )
  )


# Set-up ------------------------------------------------------------------

# Create lists of subjids who meet criteria
all_patients <- scot_data %>%
  distinct(subjid) %>%
  pull(subjid)

corna_confirmed <- scot_data %>%
  filter(corna_mbcat == "YES - Confirmed") %>%
  distinct(subjid) %>%
  pull(subjid)

corna_suspected <- scot_data %>%
  filter(corna_mbcat == "YES - Probable") %>%
  distinct(subjid) %>%
  pull(subjid)

corna_unknown <- scot_data %>%
  distinct(subjid) %>%
  filter(!subjid %in% c(corna_confirmed, corna_suspected)) %>%
  pull(subjid)

travelled <- scot_data %>%
  filter(travel_erterm == "Yes") %>%
  distinct(subjid) %>%
  pull(subjid)

died <- scot_data %>%
  filter(dsterm == "Death") %>%
  distinct(subjid) %>%
  pull(subjid)

male <- scot_data %>%
  filter(sex == "Male") %>%
  distinct(subjid) %>%
  pull(subjid)

female <- scot_data %>%
  filter(sex == "Female") %>%
  distinct(subjid) %>%
  pull(subjid)

# Breakdown of gender, age and death
# Used for population pyramid


pop_data <- scot_data %>%
  group_by(subjid) %>%
  filter(!subjid %in% travelled) %>%
  mutate(died = if_else(subjid %in% died, 1, 0)) %>%
  summarise(
    age = first(na.omit(age.factor)),
    sex = first(na.omit(sex)),
    ethnicity = first(na.omit(ethnicity)),
    pregnancy = first(na.omit(pregyn_rptestcd)),
    admission = first(na.omit(cestdat)),
    outcome = first(na.omit(dsterm))
  ) %>%
  mutate(died = case_when(outcome == "Death" ~ "Yes", TRUE ~ "No") %>% factor(levels = c("No", "Yes")))
```

*Data Extracted from CoCIN - `r format(extract_date, "%A, %d %B, %Y - %H:%M%p")`*

As of the above extract time the database contains records for:

# Can't work out how to get subsetted list to work!!
* `r length(all_patients)` patients in Scottish hospitals, `r length(died)` of those patients have died.
* `r length(corna_confirmed)` are confirmed CoVID-19 cases
* `r length(corna_suspected)` are suspected cases
* `r length(corna_unknown)` are unknown cases

* `r length(travelled)` patients had travelled in the 14 days before admission and have been excluded from all subsequent analyses
* `r nrow(pop_data %>% filter(pregnancy == "YES"))` patients are identified as being pregnant (excluding those who travelled)

## NHS Health Board summary

```{r hb_analysis, include = FALSE}
hb_summary_ex_travelled <- scot_data %>%
  mutate(HB2019Name = as_factor(HB2019Name) %>% fct_explicit_na(na_level = "Hospital code errors")) %>%
  group_by(HB2019Name) %>%
  filter(!subjid %in% travelled & !(is.na(HB2019Name))) %>%
  distinct(subjid, .keep_all = T) %>%
  summarise(
    total_cases = n(),
    has_chi = sum(!is.na(nhs_chi)),
    n_hosptials = n_distinct(hospid),
    confirmed_cases = sum(subjid %in% corna_confirmed),
    suspected_cases = sum(subjid %in% corna_suspected),
    unknown_cases = sum(subjid %in% corna_unknown),
    deaths = sum(subjid %in% died)
  )
```

```{r hb_table}
hb_summary_ex_travelled %>%
  gt() %>%
  tab_options(table.align = "left") %>% 
  tab_header(
    title = "Summary of NHS Scotland Health Boards",
    subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
      n_travelled = length(travelled)
    )
  ) %>%
  cols_align(
    align = "right"
  ) %>%
  cols_align(
    align = "left",
    columns = vars(HB2019Name)
  ) %>%
  cols_label(
    HB2019Name = "NHS Health Board",
    total_cases = "Number of cases",
    has_chi = "Number of patients with CHI number",
    n_hosptials = "Number of hospitals",
    confirmed_cases = "Number of confirmed",
    suspected_cases = "Number of suspected",
    unknown_cases = "Number where status is unknown",
    deaths = "Number of deaths"
  ) %>%
  summary_rows(
    columns = vars(total_cases, has_chi, n_hosptials, confirmed_cases, suspected_cases, unknown_cases, deaths),
    fns = list(
      Scotland = ~ sum(.)
    ), drop_trailing_zeros = TRUE
  )
```

## Age Breakdown
```{r age_analysis, include = FALSE}
# Age analysis
age_summary <- pop_data %>%
  group_by(age) %>%
  summarise(total_cases = n()) %>%
  mutate(percent = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

na_age_pct <- (nrow(pop_data %>% filter(is.na(age))) / nrow(pop_data)) * 100
```

*Age is unknown for `r na_age_pct`% of patients*
```{r age_table}
age_summary %>%
  gt() %>%
  tab_options(table.align = "left") %>% 
  tab_header(
    title = "Age breakdown",
    subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
      n_travelled = length(travelled)
    )
  ) %>%
  cols_align(
    align = "right"
  ) %>%
  cols_align(
    align = "left",
    columns = vars(age)
  ) %>%
  cols_label(
    age = "Age band",
    total_cases = "Number of cases",
    percent = "Percentage"
  )
```

## Population pyramid
Patients who have died are highlighted
```{r pop_pyramid}
# Population pyramid
ggplot(pop_data, aes(x = age, fill = sex, alpha = died)) +
  geom_bar(data = filter(pop_data, sex == "Female")) +
  geom_bar(data = filter(pop_data, sex == "Male"), aes(y = ..count.. * (-1))) +
  scale_fill_discrete("Sex at birth") +
  scale_alpha_ordinal("Died", range = c(0.3, 1)) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom") +
  ylim(-100, 100) +
  xlab("Age") +
  ylab("Count of subjects") +
  ggtitle("Population pyramid of ISARIC subjects from Scottish hospitals",
    subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
      n_travelled = length(travelled)
    )
  )
```

## Ethnicity Breakdown
```{r ethnicity_analysis, include = FALSE}
## Ethnicity analysis

ethnicity_data <- pop_data %>%
  group_by(ethnicity) %>%
  filter(!subjid %in% travelled) %>%
  summarise(total_cases = n()) %>%
  mutate(percent = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

white_ethn_pct <- (nrow(pop_data %>% filter(ethnicity == "White")) / nrow(pop_data)) * 100
na_ethn_pct <- (nrow(pop_data %>% filter(is.na(ethnicity))) / nrow(pop_data)) * 100
```
*`r white_ethn_pct`% of patients are of white ethnicity. Ethnicity is unknown for `r na_ethn_pct`% of patients.* 
```{r ethnicity_table}
ethnicity_data
```

## Onset Week
```{r onset_week_analysis, include = FALSE}
## Admission date analysis
## Analysing by onset date of first/earliest symptom
## Week 1 - 23/03/2020

admission_data <- pop_data %>%
  filter(admission <= today()) %>%
  mutate(
    admissionweek =
      case_when(
        admission < "2020-03-30" ~ "Week 1",
        admission < "2020-04-06" ~ "Week 2",
        admission < "2020-04-13" ~ "Week 3",
        admission < "2020-04-20" ~ "Week 4",
        admission < "2020-04-27" ~ "Week 5",
        admission < "2020-05-04" ~ "Week 6",
        TRUE ~ "NA"
      )
  ) %>%
  group_by(admissionweek) %>%
  summarise(total_cases = n()) %>%
  mutate(percent = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

week1_date <- "March 23 2020"
latest_week <- max(admission_data %>% pull(admissionweek))
```

This analysis using onset date of first/earliest symptom (some patients were in hospital before the start of the pandemic). Week 1 has been selected as the earliest week in which a patient showed symptoms – `r week1_date`, `r latest_week` is the current week.

```{r onset_week_table}
admission_data
```
