---
title: "Scotland Summary statistics"
author: "Alice Byres, Nicole Jarvie and James McMahon"
date: "`r lubridate::stamp_date('Thursday, 30 April, 2020')(lubridate::today())`"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(lubridate)
library(finalfit)
library(RCurl)
library(tidylog)
library(Hmisc)


if (!exists("scot_data")) { # Use this with rmarkdown::render("scotland_summary.Rmd") for testing
  source("code/00 - Get Scottish data.R")
  source("code/01 - Prep data.R")
}
```
*Data Extracted from CoCIN - (save date and time for extract to be used here)*

As of the above extract time the database contains records for:

* `r "XX"` patients in Scottish hospitals
* `r "XX"` of those patients have died
* Of these patients

  + `r "XX"` are confirmed CoVID-19 cases
  + `r "XX"` are suspected cases
  + `r "XX"` are unknown cases
* `r "XX"` patients are identified as being pregnant
* `r "XX"` patients had travelled in the 14 days before admission and have been excluded from the analyses below

## NHS Health Board summary

```{r hb_table}

```

## Age Breakdown
*Age is unknown for `r "X.X"`% of patients*
```{r age_table}

```

## Population pyramid
Patients who have died are highlighted
```{r pop_pyramid}

```

## Ethnicity Breakdown
*`r "X.X"`% of patients are of white ethnicity. Ethnicity is unknown for `r "X.X"`% of patients.* 
```{r ethnicity_table}

```

## Onset Week
This analysis using onset date of first/earliest symptom (some patients were in hospital before the start of the pandemic). Week 1 has been selected as the earliest week in which a patient showed symptoms – `r "DATE"`, week `r "n"` is the current week.

```{r onset_week_table}

```
