---
title: "Scotland Summary statistics"
author: "Alice Byers, Nicole Jarvie and James McMahon"
date: "`r format(Sys.Date(), '%A, %d %B, %Y') `"
output: word_document
---
<!-- Note that stored_ccp_token.txt must be created with your API token before this report will work. There is a block of code in 00 which will create it easily -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(lubridate)
library(scales)
library(finalfit)
library(RCurl)
library(Hmisc)
library(ggplot2)
library(openxlsx)
library(gt)
library(ggplot2)


if (!exists("scot_data")) { # Use this with rmarkdown::render("scotland_summary.Rmd") for testing
  source("code/00 - Get Scottish data.R")
  source("code/01 - Prep data.R")
}
```

```{r data_prep, include = FALSE}
# Fixes -------------------------------------------------------------------
# Change the age groupings
scot_data <- scot_data %>%
  mutate(
    age.factor = case_when(
      age < 17 ~ "<17",
      age < 30 ~ "17-29",
      age < 40 ~ "30-39",
      age < 50 ~ "40-49",
      age < 60 ~ "50-59",
      age < 70 ~ "60-69",
      age < 80 ~ "70-79",
      is.na(age) ~ NA_character_,
      TRUE ~ "80+"
    )
  )


# Set-up ------------------------------------------------------------------

# Create lists of subjids who meet criteria
all_patients <- scot_data %>%
  distinct(subjid) %>%
  pull(subjid)

corna_confirmed <- scot_data %>%
  filter(corna_mbcat == "YES - Confirmed") %>%
  distinct(subjid) %>%
  pull(subjid)

corna_suspected <- scot_data %>%
  filter(corna_mbcat == "YES - Probable") %>%
  distinct(subjid) %>%
  pull(subjid)

corna_unknown <- scot_data %>%
  distinct(subjid) %>%
  filter(!subjid %in% c(corna_confirmed, corna_suspected)) %>%
  pull(subjid)

travelled <- scot_data %>%
  filter(travel_erterm == "Yes") %>%
  distinct(subjid) %>%
  pull(subjid)

died <- scot_data %>%
  filter(dsterm == "Death") %>%
  distinct(subjid) %>%
  pull(subjid)

male <- scot_data %>%
  filter(sex == "Male") %>%
  distinct(subjid) %>%
  pull(subjid)

female <- scot_data %>%
  filter(sex == "Female") %>%
  distinct(subjid) %>%
  pull(subjid)

# Breakdown of gender, age and death
# Used for population pyramid
pop_data <- scot_data %>%
  group_by(subjid) %>%
  filter(!subjid %in% travelled) %>%
  mutate(died = if_else(subjid %in% died, 1, 0)) %>%
  summarise(
    age = first(na.omit(age.factor)),
    sex = first(na.omit(sex)),
    ethnicity = first(na.omit(ethnicity)),
    pregnancy = first(na.omit(pregyn_rptestcd)),
    symptom_onset = first(na.omit(cestdat)),
    admission = first(na.omit(hostdat)),
    outcome = first(na.omit(dsterm))
  ) %>%
  mutate(
    died = case_when(outcome == "Death" ~ "Yes", TRUE ~ "No") %>%
      factor(levels = c("No", "Yes")),
    status = case_when(
      outcome %in% c("Death", "Palliative discharge") ~ "Died",
      outcome %in% c("Discharged alive") ~ "Discharged",
      TRUE ~ "On-going care"
    ) %>%
      factor(levels = c("Discharged", "On-going care", "Died"))
  )
```

*Data Extracted from CoCIN - `r format(extract_date, "%A, %d %B, %Y - %H:%M%p")`*

As of the above extract time the database contains records for:

<!---Can't work out how to get subsetted list to work!!-->
* `r length(all_patients)` patients in Scottish hospitals, `r length(died)` of those patients have died.
* `r length(corna_confirmed)` are confirmed CoVID-19 cases
* `r length(corna_suspected)` are suspected cases
* `r length(corna_unknown)` are unknown cases

* `r length(travelled)` patients had travelled in the 14 days before admission and have been excluded from all subsequent analyses
* `r nrow(pop_data %>% filter(pregnancy == "YES"))` patients are identified as being pregnant (excluding those who travelled)

## NHS Health Board summary


```{r hb_analysis, include = FALSE}
hb_summary_ex_travelled <- scot_data %>%
  mutate(HB2019Name = as_factor(HB2019Name) %>% fct_explicit_na(na_level = "Hospital code errors")) %>%
  group_by(HB2019Name) %>%
  filter(!subjid %in% travelled & !(is.na(HB2019Name))) %>%
  distinct(subjid, .keep_all = T) %>%
  summarise(
    total_cases = n(),
    has_chi = sum(!is.na(nhs_chi)),
    n_hosptials = n_distinct(hospid),
    confirmed_cases = sum(subjid %in% corna_confirmed),
    suspected_cases = sum(subjid %in% corna_suspected),
    unknown_cases = sum(subjid %in% corna_unknown),
    deaths = sum(subjid %in% died)
  ) 
```


```{r hb_table}
hb_summary_ex_travelled %>%
  pander::pander(split.table = Inf)

# hb_summary_ex_travelled %>%
#   gt() %>%
#   tab_options(table.align = "left") %>%
#   tab_header(
#     title = "Summary of NHS Scotland Health Boards",
#     subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
#       n_travelled = length(travelled)
#     )
#   ) %>%
#   cols_align(
#     align = "right"
#   ) %>%
#   cols_align(
#     align = "left",
#     columns = vars(HB2019Name)
#   ) %>%
#   cols_label(
#     HB2019Name = "NHS Health Board",
#     total_cases = "Number of cases",
#     has_chi = "Number of patients with CHI number",
#     n_hosptials = "Number of hospitals",
#     confirmed_cases = "Number of confirmed",
#     suspected_cases = "Number of suspected",
#     unknown_cases = "Number where status is unknown",
#     deaths = "Number of deaths"
#   ) %>%
#   summary_rows(
#     columns = vars(total_cases, has_chi, n_hosptials, confirmed_cases, suspected_cases, unknown_cases, deaths),
#     fns = list(
#       Scotland = ~ sum(.)
#     ), drop_trailing_zeros = TRUE
#   )
```

## Age/Sex Breakdown
```{r age_sex_analysis, include = FALSE}
# Age/sex analysis
age_sex_summary <- pop_data %>%
  group_by(age) %>%
  summarise(total_cases = n(),
            males = sum(sex == "Male", na.rm = TRUE),
            females = sum(sex == "Female", na.rm = TRUE)) %>%
  mutate(percent = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

na_age_pct <- (nrow(pop_data %>% filter(is.na(age))) / nrow(pop_data)) * 100
na_sex_pct <- (nrow(pop_data %>% filter(is.na(sex))) / nrow(pop_data)) * 100
```

*Age is unknown for `r number(na_age_pct, accuracy = 0.1)`% of patients. Sex is unknown for `r number(na_sex_pct, accuracy = 0.1)`% of patients.*
```{r age_sex_table}
age_sex_summary %>%
  pander::pander()

# age_sex_summary %>%
#   gt() %>%
#   tab_options(table.align = "left") %>%
#   tab_header(
#     title = "Age/Sex breakdown",
#     subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
#       n_travelled = length(travelled)
#     )
#   ) %>%
#   cols_align(
#     align = "right"
#   ) %>%
#   cols_align(
#     align = "left",
#     columns = vars(age)
#   ) %>%
#   cols_label(
#     age = "Age band",
#     total_cases = "Number of cases",
#     males = "Number of Males",
#     females = "Number of Females",
#     percent = "Percentage"
#   ) %>% 
#   summary_rows(
#     columns = vars(total_cases, males, females),
#     fns = list(
#       Total = ~ sum(.)
#     ), drop_trailing_zeros = TRUE
#   )
```

## Population pyramid
Patients who have died are highlighted
```{r pop_pyramid}
# Population pyramid
# James' old code
# max_y <- pop_data %>% count(age, sex) %>% pull(n) %>% max()
# 
# ggplot(pop_data, aes(x = age, fill = sex, alpha = died)) +
#   geom_bar(data = filter(pop_data, sex == "Female")) +
#   geom_bar(data = filter(pop_data, sex == "Male"), aes(y = ..count.. * (-1))) +
#   scale_fill_discrete("Sex at birth") +
#   scale_alpha_ordinal("Died", range = c(0.3, 1)) +
#   scale_y_continuous(n.breaks = 8, labels = abs, limits = c(-max_y, max_y)) +
#   coord_flip() +
#   theme_minimal() +
#   theme(legend.position = "bottom") +
#   xlab("Age band") +
#   ylab("Count of patients") 

# New code, mostly stolen from UoE (actually quite similar)
max_y <- pop_data %>% 
  count(age, sex) %>% 
  pull(max(n))

pyramid_text = pop_data %>% 
  count(sex) %>% 
  filter(sex != "Not specified") %>% 
  mutate(total_n = sum(n),
         prop = scales::percent(n/total_n),
         label = paste0(n, "/", total_n, " (", prop, ")"))

pop_data %>%
  ggplot(aes(x = age, fill = status)) +
  geom_bar(data = subset(pop_data, sex == "Female")) +
  geom_bar(data = subset(pop_data, sex == "Male"), aes(y = ..count.. * (-1))) +
  geom_hline(yintercept = 0, colour = "navyblue") +
  scale_y_continuous(n.breaks = 8, labels = abs, limits = c(-max(max_y), max(max_y))) +
  # scale_y_continuous(
  #   breaks = seq(-(ceiling(max(max_y) / 100) * 100), ceiling(max(max_y) / 100) * 100, 50),
  #   labels = abs(seq(-(ceiling(max(max_y) / 100) * 100), ceiling(max(max_y) / 100) * 100, 50)),
  #   limits = c(-ceiling(max(max_y) / 100) * 100, ceiling(max(max_y) / 100) * 100)
  # ) +
  # expand_limits(y = c(-pyramid_max, pyramid_max)) +
  scale_fill_brewer("", palette = "Blues") +
  coord_flip() +
  annotate("text", x = "17-29", y = -ceiling(max(max_y) / 100) * 70, label = "Males", size = 5, hjust = 0, vjust = 0.2) +
  annotate("text", x = "17-29", y = ceiling(max(max_y) / 100) * 70, label = "Females", size = 5, hjust = 1, vjust = 0.2) +
  annotate("text", x = "<17", y = -ceiling(max(max_y) / 100) * 70, label = pyramid_text$label[1], size = 5, hjust = 0) +
  annotate("text", x = "<17", y = ceiling(max(max_y) / 100) * 70, label = pyramid_text$label[2], size = 5, hjust = 1) +
  labs(x = "Age (years)", y = "Number of patients") +
  theme_classic() +
  theme(legend.position = "top") +
  guides(fill = guide_legend(nrow = 1)) +
  ggtitle(
    "Patients with outcome stratified by age, and sex"
  )

# +
#   ggtitle("Population pyramid of ISARIC subjects from Scottish hospitals",
#     subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
#       n_travelled = length(travelled)
#     )
#   )
```

## Ethnicity Breakdown
```{r ethnicity_analysis, include = FALSE}
## Ethnicity analysis

ethnicity_data <- pop_data %>%
  group_by(ethnicity) %>%
  filter(!subjid %in% travelled) %>%
  summarise(total_cases = n()) %>%
  mutate(percent = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

white_ethn_pct <- (nrow(pop_data %>% filter(ethnicity == "White")) / nrow(pop_data)) * 100
na_ethn_pct <- (nrow(pop_data %>% filter(is.na(ethnicity))) / nrow(pop_data)) * 100
```
*`r number(white_ethn_pct, accuracy = 0.1)`% of patients are of white ethnicity. Ethnicity is unknown for `r number(na_ethn_pct, accuracy = 0.1)`% of patients.* 
```{r ethnicity_table}
ethnicity_data%>%
  pander::pander()

# ethnicity_data %>%
#   gt() %>%
#   tab_options(table.align = "left") %>%
#   tab_header(
#     title = "Ethnicity breakdown",
#     subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
#       n_travelled = length(travelled)
#     )
#   ) %>%
#   cols_align(
#     align = "right"
#   ) %>%
#   cols_align(
#     align = "left",
#     columns = vars(ethnicity)
#   ) %>%
#   cols_label(
#     ethnicity = "Ethnicity",
#     total_cases = "Number of cases",
#     percent = "Percentage"
#   )
```

## Onset Week
```{r onset_week_analysis, include = FALSE}
## Onset of symptoms date analysis
## Analysing by onset date of first/earliest symptom
## Week 1 - 23/03/2020

admission_data <- pop_data %>%
  filter(between(admission, dmy("23/03/2020"), today())) %>%
  mutate(
    week = floor_date(admission, unit = "week", week_start = 1),
    iso_week = isoweek(admission)
  ) %>%
  group_by(week, iso_week) %>%
  summarise(total_admissions = n()) %>%
  ungroup() %>% 
  mutate(percent_admissions = paste(round(total_admissions / sum(total_admissions) * 100, 1), "%"))

symptom_data <- pop_data %>%
  filter(between(symptom_onset, dmy("23/03/2020"), today())) %>%
  mutate(
    week = floor_date(symptom_onset, unit = "week", week_start = 1),
    iso_week = isoweek(symptom_onset)
  ) %>%
  group_by(week, iso_week) %>%
  summarise(total_cases = n()) %>%
  ungroup() %>% 
  mutate(percent_cases = paste(round(total_cases / sum(total_cases) * 100, 1), "%"))

onset_data <- left_join(symptom_data, admission_data)

week1_date <- "March 23 2020"
latest_week <- max(admission_data %>% pull(week))
```

This analysis using onset date of first/earliest symptom (some patients were in hospital before the start of the pandemic). The earliest week in which a patient showed symptoms – `r week1_date` (ISO week `r isoweek(mdy(week1_date))`), `r latest_week` () (ISO week `r isoweek(mdy(latest_week))`), is the current week.

```{r onset_week_table}
onset_data %>%
  pander::pander(split.table = Inf)

# admission_data %>%
#   gt() %>%
#   tab_options(table.align = "left") %>%
#   tab_header(
#     title = "Week of Admission",
#     subtitle = str_glue("Excludes {n_travelled} who travelled in the 14 days prior to admission",
#       n_travelled = length(travelled)
#     )
#   ) %>%
#   cols_align(
#     align = "right"
#   ) %>%
#   cols_align(
#     align = "left",
#     columns = vars(admissionweek)
#   ) %>%
#   cols_label(
#     admissionweek = "Week of admission",
#     total_cases = "Number of cases",
#     percent = "Percentage"
#   )
```
